---
title: "Assignment_3"
author: "Hana Pham"
date: "2/14/2022" 
output: html_document
---

```{r setup, include =FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(maps)
library(mapproj)
library(ggplot2)
```

## Introduction & Summary Info: 

This report represents data on the inequality within the United States incarceration system. The main goal is to understand how different incarceration measures vary by race. I've chosen to compare the prevalent minority communities (from ages 15 to 64) Black, Latinx, and Asian American & Pacific Islander to the White population. For each, I found the maximum value for the prisoner population, for instance `max_black`. I then evaluated the difference of these maximum values with the maximum value for White prisoners like I calculated with `diff_latinx_white`. The purpose was to portray the drastic difference minorities face, compared to white people, in terms of the U.S. prison system. I also wanted to show the variation within the minority populations themselves. My findings painted a picture that I had anticipated. The values I calculated related to prisoner populations of Black, Latinx, and Asian American & Pacific Islanders showed the disproportionate incarceration of people of color. Another value I reported on was the county with the highest Black prisoner population. With `highest_county_black`, I found that New York County had the highest number of black prisoners. On the other hand, `lowest_county` depicted that Logan County had the lowest number of black prisoners. I followed that with `highest_county_white`: the value of the maximum number of White prisoners in that same county, and I could see that these two values were not different from one another. Finally, I wanted to focus on the ratio of black prisoners to the total population in comparison to the white prisoner ratio. My `final_ratio` variable showed that black prisoners compiled a large part of the total prisoners. Overall, each variable calculated a value that confirmed an all too popularly known fact: that there is a disproportionate incarceration of people of color within the United States justice system. 

```{r, echo = FALSE}
source("../docs/analysis.R")
library(dplyr)
library(tidyverse)
library(ggplot2)
```

## *Charts*

#### _*Trends Over Time Chart*_
This trend chart was the focal point of my questions while looking at the data set of incarceration trends. I was mainly comparing white prisoner populations to minority prisoner populations. This line graph compares black and white prisoner populations from ages 15 to 64. The trajectory of the white prisoners is steadily declining from the year 2000 towards 2010. However, the black prisoner population is _steadily holding._ There is not even a slight change. This visualization draws attention to the *extreme* lack of equality within the prison system, and it portrays the disparities minorities face in the justice system.

```{r trend_chart, echo=FALSE}
source("../docs/analysis.R")
ggplot(trend_of_white_vs_black_population) + 
      geom_smooth(mapping = aes(x = year, y = population, fill = race)) +
      labs(title = "Trends of White vs. Black Prisoner Populations", 
           x = "Year", y= "Population Number", fill = "Race") 
```

#### _*Variable Comparison Chart*_ 
This scatterplot takes another minority population, Latinx, and compares it with the white prisoner population. I included a line with the scatterplot to show the sudden upward incline of the Latinx prisoner population compared to white prisoners. This scatterplot shows a noticeable rise in the population of Latinx in jails. However, along the x-axis, you can see that the white prisoner population stays relatively lower than Latinx. This data comparison is another clear indication of the disproportionate incarceration of minority communities. This visualization also includes data on the concentration of the groups among the U.S. regions. The purpose of that was to show the distribution of prisoner populations. It shows that the West experienced the most frequent Latinx prisoners, and the Northeast region.

``` {r variable_comparison, echo = FALSE}
source("../docs/analysis.R")
ggplot(white_vs_latinx_population) + 
    geom_point(mapping = aes(x= white_pop_15to64, y = latinx_pop_15to64, color = region)) + 
    labs(title = "Latinx vs. White Prisoner Population",
         x = "White Population", y ="Latinx Population", color = "Region") +
    geom_smooth(mapping = aes(x = white_pop_15to64, y = latinx_pop_15to64))
```

#### _*Map*_ 
The main message I wanted to portray with this map was to represent the widespread problem of high prisoner populations composed of black people. This map shows the overall distribution and proves how prevalent this inequality is in the United States. We can notice from this visualization that there is a heavy concentration of black prisoners on the West Coast of the U.S. There is scarcely any in the middle of the map, and there are small concentrated clusters in the South and East Coast, too. This takeaway is not surprising as these locations suffer from racial disparities. 

``` {r blank_theme, map, echo = FALSE}
source("../docs/analysis.R")
library(maps)
library(mapproj)
ggplot(map_data1) +
    geom_polygon(
      mapping = aes(x = long, y = lat, group = group, fill = black_pop_15to64),
      color = "gray",
      size = 0.3) +
    coord_map() + 
    scale_fill_continuous(limits = c(0,max(map_data1$black_pop_15to64)), na.value = "white", low = "yellow", high = "red") +
    labs(title = "Black Prisoner Population in the US", fill = "Black Prisoner Population") +
    blank_theme 
```

